Assessing Power of the California Microhap Baseline for GSI (and calculating Fst)

### 2024-04-09 002-GSI_and_Fst.Rmd
Add GSI table with the LFAR

```{r input_bypass, include=NEG_USING_SMK, eval=NEG_USING_SMK}
# inputs:
input_list <- list(
  #final_baseline = "data/cali-chinook-baseline.rds",
  final_baseline = "data/subRoSA_baseline_rubias_FEB2024_w_mixture_revised.csv",
  pop_labels = "inputs/reference-collection-names.csv",
  map_notations = "inputs/map-notations.tsv"
)
# outputs:
output_list <- list(
  assig_table = "results/GSI_and_Fst/assignment-table-full-baseline.pdf",
  fst_table = "results/GSI_and_Fst/full-baseline-but-no-lfar-Fst-table.pdf",
  assig_table_lfar_all = "results/GSI_and_Fst/assignment-table-full-baseline-plus-lfar.pdf",
  assig_table_lfar_no_miss = "results/GSI_and_Fst/assignment-table-full-baseline-plus-lfar-and-everyone-has-the-lfar.pdf",
  tex_table = "tex/inputs/samples-table.tex",
  gsi_fst_fig = "results/GSI_and_Fst/gsi_and_fst_fig.pdf",
  gsi_fst_fig_crop = "results/GSI_and_Fst/gsi_and_fst_fig-crop.pdf",
  gsi_fst_fig_tex = "tex/images/gsi_and_fst_fig-crop.pdf"
)
# you can create the necessary output directories like this:
dump <- lapply(output_list, function(x)
  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
)
```

@@

@@

>>>>

```{r input_bypass, include=NEG_USING_SMK, eval=NEG_USING_SMK}
# inputs:
input_list <- list(
  #final_baseline = "data/cali-chinook-baseline.rds",
  final_baseline = "data/subRoSA_baseline_rubias_FEB2024_w_mixture_revised.csv",
  lfar_genos = "data/metagenos_hap2col_lfar_all_samps.csv",
  pop_labels = "inputs/reference-collection-names.csv",
  map_notations = "inputs/map-notations.tsv"
)
# outputs:
output_list <- list(
  assig_table = "results/GSI_and_Fst/assignment-table-full-baseline.pdf",
  fst_table = "results/GSI_and_Fst/full-baseline-but-no-lfar-Fst-table.pdf",
  assig_table_lfar_all = "results/GSI_and_Fst/assignment-table-full-baseline-plus-lfar.pdf",
  tex_table = "tex/inputs/samples-table.tex",
  gsi_fst_fig = "results/GSI_and_Fst/gsi_and_fst_fig.pdf",
  gsi_fst_fig_crop = "results/GSI_and_Fst/gsi_and_fst_fig-crop.pdf",
  gsi_fst_fig_tex = "tex/images/gsi_and_fst_fig-crop.pdf"
)
# you can create the necessary output directories like this:
dump <- lapply(output_list, function(x)
  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
)
```

@@

lfar_genos <- read_csv(input_list$lfar_genos)

@@

# Now, we repeat that while adding the LFAR markers in there

We do these as an additional step to get a sense for how much
improvement we get from adding the LFAR in there.

First we need to join the LFARs on there:
```{r}
# get just the 3 LFAR markers we are using:
lfar_columns <- paste(
  rep(c(
  "NC_037130.1:1062935-1063235",
  "NC_037130.1:828619-828919",
  "NC_037130.1:864908-865208"
  ), each = 2),
  rep(c(1,2), 3),
  sep = "_"
)

lfar3genos <- lfar_genos[c("Sample_ID", lfar_columns)]

base_with_lfar <- full_base2 %>%
  left_join(lfar3genos, by = join_by(indiv == Sample_ID))
```

Then, all the step from above:
```{r}
# Do self-assignment:
full_plus_lfar_sa <- self_assign(base_with_lfar, gen_start_col = 5)

top_ass_w_lfar <- full_plus_lfar_sa %>%
  group_by(indiv, collection, repunit, inferred_repunit) %>%
  mutate(repu_sclike = sum(scaled_likelihood)) %>%
  group_by(indiv) %>%
  filter(repu_sclike == max(repu_sclike)) %>%
  arrange(indiv, desc(scaled_likelihood)) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(coll_f = factor(collection, levels = collection_order))

investigate the ones that are in the wrong repunit, and look at their
scaled_repu_likelihoods, the others.
```{r}
ta_corr_lfar <- top_ass_w_lfar %>%
  filter(repunit == inferred_repunit)
ta_wrong_lfar <- top_ass_w_lfar %>%
  filter(repunit != inferred_repunit) %>%
  arrange(collection, repu_sclike) %>%
  group_by(collection) %>%
  mutate(y = 190 + (70/9) * 1:n())

ggplot() +
  geom_histogram(
    data = ta_corr_lfar,
    mapping = aes(x = repu_sclike, fill = inferred_repunit)
  ) +
  geom_point(
    data = ta_wrong_lfar,
    mapping = aes(x = repu_sclike, y = y, fill = inferred_repunit),
    size = 3,
    shape = 21
  ) +
  facet_wrap(~coll_f, ncol = 3) +
  scale_fill_manual(values = repunit_colors)
```

Filter out the ones that look like clear mis-labelling and then
count up the mis-assigned fish:
```{r}
tafilt_w_lfar <- top_ass_w_lfar %>%
  filter(!(repunit != inferred_repunit & repu_sclike > 0.99))
```

Here we calculate the number of correctly and incorrectly assigned fish:
```{r}
tafilt_w_lfar %>%
  mutate(correct = repunit == inferred_repunit) %>%
  count(correct) %>%
  mutate(fract = n / sum(n))
```

Having the LFAR in there improves things. Good.

Make a table of those:
```{r}

X2_lfar <- tafilt_w_lfar %>%
  mutate(
    collection_f = factor(collection, levels = collection_order),
    inferred_coll_f = factor(inferred_collection, levels = collection_order)
  ) %>%
  count(collection_f, inferred_coll_f, .drop = FALSE) %>%
  taf_prep()

Rubias_for_later_lfar <- X2_lfar # keeping this for later...

# get the result
TAF2_lfar <- table_as_figure(
  X = X2_lfar,
  RC_groups = RC_groups,
  external_colors = run_time_colors,
  internal_colors = repunit_colors,
  plot_margins = c(1, 0.1, 0.7, 0.1)
)

# and plot it
g_lfar <- plot_grid(
  TAF2_lfar$full_plot,
  plot_grid(
    get_legend(TAF2_lfar$for_external_legend),
    get_legend(TAF2_lfar$for_internal_legend),
    nrow = 1
  ),
  nrow = 2,
  rel_heights = c(7,3)
)

g_lfar
```

Adding the LFAR definitely improves the LFAR assignment, though
it still does not get it perfect.
```{r}
ggsave(g_lfar, filename = output_list$assig_table_lfar_all, width = 10, height = 12)
```
